|

### 2021-12-16 07-Chp-Triangulation.Rmd
End of day:

This approach represents a significant step forward from the current practice of synthesising evidence from only one type of study (e.g. randomised controlled trials (RCTs), non-randomised studies of exposures (NRSE) or interventions (NRSI)). The most common implementation of this method to date has been in the form of _qualitative triangulation_ - that is the identification and narrative comparison of diverse evidence sources with respect to their underlying bias structures.[@lawlor2016]

However, qualitative triangulation faces issues at scale. There is a need to include all evidence to avoid potential confirmation bias,[@dubroff2018] yet as the number of evidence sources increases, it is not possible to narratively compare and contrast across multiple results in any meaningful way. This fact is illustrated by previous exemplars of triangulation only considering a small number of individual results.[@ference2014; @lawlor2016] To overcome this limitation, some attempts have considered the output of a meta-analysis of similarly-designed studies to be a single source of evidence, and have assessed bias in relation to this summary result. While this approach keeps the number of individual sources of evidence manageable, it loses useful information on the specific biases inherent to each result contributing to the summary estimate. For example, while it may be true that all non-randomised studies share some minimal level of bias, competing extents and directions of bias in individual results may be masked by the use of summary estimates. In addition, previous attempts at qualitative triangulation have not systematically assessed the indirectness of the results, namely the differences between the question of interest and the question addressed by a study.

@@

Meta-analysis of lipid levels in the systematic review found extremely weak evidence for an association of any lipid fraction with Alzheimer's disease (TC - `r obsLipids$AD_TC$estimate`; LDL - `r obsLipids$AD_LDL$estimate`; HDL - `r obsLipids$AD_HDL$estimate`; TG - `r obsLipids$AD_TG$estimate`). Similarly, meta-analysis of studies examining a binary hypercholesterolemia provided weak evidence for an association of this risk factor with Alzheimer's disease (`r obsHyperchol$AD$estimate`). A meta-analysis of MR analyses of genetic instruments that mimic statin use found weak evidence for an association of lipid lowering with Alzheimer's disease (`r mrStatin$AD$estimate`).

@@

Finally, only one of the three cohorts included in the IPD analysis (Whitehall II) contained data on Alzheimer's disease. However, a previously published analysis (Tynkkynen _et al._[@tynkkynen2018]) examining the association of lipid levels with Alzheimer's disease in this cohort was identified by the systenatic review. The results of this previous analysis are incorporated into the results of the meta-analysis reported above, and so the IPD analysis provided no new information on this outcome.

@@

These causal questions were chosen for several reasons. Firstly, the work presented in the previous chapters had identified some evidence for an association in these exposure/outcome pairs. Secondly, given the long prodromal period between the onset of physiological changes in the brain and presentation of dementia symptoms, it seems likely that conditions during the mid-life period are particularly important. Finally, examining the causal impact of a lipid fraction on a specific outcome (e.g. Alzheimer's disease), rather than a composite outcome (e.g. all-cause dementia), focuses on an single causal pathway rather than including several conditions in the outcome definition that likely have very different mechanisms of disease. In addition, previous work presented in this thesis illustrated that the component conditions (in this case Alzheimer's disease and vascular dementia) are likely to be subject to very different confounding structures (see Section \@ref(cprd-confounding-by-ind)), and the use of a composite all-cause outcome in the triangulation framework may mask this.

@@

Once the causal question of interest had been defined, relevant individual results were obtained from the data sources described in Section \@ref(tri-data-sources). In a broad sense, this meant extracting results that examined the relationship between the exposure/outcome pair, either directly (e.g. non-randomised studies of LDL-c levels) or indirectly (e.g. non-randomised studies of statins, or MR studies using genetic instruments for LDL-c levels).

@@

&nbsp;<!----------------------------------------------------------------------->  
(ref:robLevelsMapping-caption) __Mapping of risk-of-bias judgements across assessment tools__ The different acceptable levels of bias across the risk-of-bias assessment tools used were mapped to a harmonised set. Note that no mapping for "Critical" risk of bias was performed, as best practice recommends that studies at critical risk of bias are excluded from quantitative syntheses.

@@

<!-- QUESTION text -->

@@

<!-- TODO Need to include some text here on examples of each: e.g. immortal time bias, or  -->

@@

One key feature of the domain-based risk-of-bias tools is that the domains are considered interchangeable - i.e. a "high" judgement in one domain is equivalent to a "high" judgement in any other. As such, for this analysis, I assumed that the distributions assigned to each extent of bias were identical across all values of $j$, that is, were identical for all domains in the risk-of-bias assessment tool. To illustrate for additive biases:

@@

(ref:priorsAdd-caption) __Prior distributions mapped to different extents of bias__ - Prior values assigned to additive and proportional biases, defined as log-normal distributions with mean $\mu$ and variance $\sigma^2$ using the notation $N(\mu,\sigma^{2})$. For additive biases, two scenarios were defined (see main text). Note that proportional biases are defined on the log scale for ease of calculation, but are exponentiated prior to their use for adjustment.

@@

Where the direction of bias in a domain was unpredictable for a given result, the mean of the prior distribution for that domain was set to 0 but the appropriate variance for the recorded extent of bias was retained (e.g. for an unpredictable "High" additive bias, the distribution would be $N(0,0.1)$ under Scenario 1). In other words, adjusting for bias in a domain with an unpredictable direction of bias has no impact on the position of the effect estimate but does increase the uncertainty around it.

Under the  method reported in Turner _et al._,[@turner2009], total additive bias (denoted using subscript $\delta$) in result $i$ is then described by mean $\mu_{i\delta}$ and variance $\sigma_{i\delta}^{2}$, which are simply the sum of the mean and variance of all additive biases in the $i$th result. Similarly, the total proportional bias (denoted using subsript $\beta$) in the $i$th result is described by mean $\mu_{i\beta}$ and variance ${\sigma_{i\beta}}^2$. This are defined as

@@

where $\alpha_i$ and ${\psi_i}^2$  are the the sum of mean and variance values for all proportional biases in the $i$th result.

@@

Using the the method reported in Turner _et al._,[@turner2009] once total additive and proportional bias and indirectness are calculated, they are used to define an adjusted estimate for each result included in the synthesis.

@@

<!-- REVIEW YBS - I may have got confused but when you say favour the intervention I am assuming you mean a “protective effect” e.g. OR <1. But if there is differential misclassification so that statin use increases diagnosis of vascular dementia that would show a deleterious effect under the null hence shift to right e.g. OR >1. Perhaps I have misunderstood favour as < risk when in fact you mean > risk. Can you clarify?

@@

From the data sources identified in Section \@ref(tri-data-sources), I identified 7 results relevant to the causal effect of triglycerides at midlife on vascular dementia. Four were identified via the systematic review,`r tg_vad_citations` while Chapter \@ref(cprd-analysis-heading) provided two additional estimates from unanalysed datasets (CaPS and Whitehall II), and Chapter \@ref(ipd-heading) provides an estimate for the effect of fibrates, a treatment for hypertriglyceridaemia. The extent, type and direction of predicted bias for each result can be seen in the bias-direction plot presented in Figure \@ref(fig:tgVadBiasDirection), stratified by study design.

@@

A further issue in this regard is that the rigour of the tools may not be equivalent. This issue can be illustrated by the Ostergaard _et al._ study[@ostergaard2015] having low risk of bias across all domains and thus receiving a substantial proportion of the weight in the first case study (Figure \@ref(fig:fpLdlAd)). While the RoB suite of tools (RoB 2.0, ROBINS-I, ROBINS-E) were developed by an expert working group and have been widely used, the Mamluk _et al._ tool[@mamluk2020] used to assess risk of bias in Mendelian randomisation studies was designed by the authors for use in their own review. If the Mamluk tool is failing to adequately assess bias in Mendelian randomisation studies, then any analysis based on its assessments is likely to be subject to bias. This observation is particularly problematic given the default position of the ROBINS-I and ROBINS-E tool to require a "Moderate" judgement in the confounding domain. Under these conditions, it is possible that a well-performed cohort study (low risk of bias across everything other that bias due to confounding) will be down-weighted versus a potentially poor MR study.

@@

The core strength of this analysis is that it is based on a comprehensive systematic review, supplemented by two additional primary studies to fill in the identified evidence gaps. In general, triangulation of any sort should be considered a natural extension of evidence synthesis, and therefore should follow best practice in relation to finding and critically assessing all relevant information. Additionally, the approach presented here also builds on recent developments in bias assessment to "explode out" the component results of a meta-analysis, and consider the effects of bias/indirectness in each separately.

@@

<!-- REVIEW Good to mention direct adjustment for bias, e.g. Greenlands writings -->

@@

* In this chapter, I narratively described the evidence identified or produced in previous chapters, and highlighted the issues with a qualitative approach to triangulation when there are several sources of evidence.

@@

* I discussed the limitations of this proposed method, given the absence of empirically-based priors for the impact of bias/indirectness, and proposed future research (i.e. meta-epidemiological studies) that would address them.

>>>>

This approach represents a significant step forward from the current practice of synthesising evidence from only one type of study (e.g., randomised controlled trials (RCTs), non-randomised studies of exposures (NRSE) or interventions (NRSI)). The most common implementation of this method to date has been in the form of _qualitative triangulation_ - that is the identification and narrative comparison of diverse evidence sources with respect to their underlying bias structures.[@lawlor2016]

However, qualitative triangulation faces issues at scale. There is a need to include all evidence to avoid potential confirmation bias,[@dubroff2018] yet as the number of evidence sources increases, it is not possible to narratively compare and contrast across multiple results in any meaningful way. This fact is illustrated by previous exemplars of triangulation only considering a small number of individual results.[@ference2014; @lawlor2016] To overcome this limitation, some attempts have considered the output of a meta-analysis of similarly designed studies to be a single source of evidence, and have assessed bias in relation to this summary result. While this approach keeps the number of individual sources of evidence manageable, it loses useful information on the specific biases inherent to each result contributing to the summary estimate. For example, while it may be true that all non-randomised studies share some minimal level of bias, competing extents and directions of bias in individual results may be masked by the use of summary estimates. In addition, previous attempts at qualitative triangulation have not systematically assessed the indirectness of the results, namely the differences between the question of interest and the question addressed by a study.

@@

Meta-analysis of lipid levels in the systematic review found extremely weak evidence for an association of any lipid fraction with Alzheimer's disease (in the systematic review: TC - `r obsLipids$AD_TC$estimate`; LDL - `r obsLipids$AD_LDL$estimate`; HDL - `r obsLipids$AD_HDL$estimate`; TG - `r obsLipids$AD_TG$estimate`). Similarly, meta-analysis of studies examining a binary hypercholesterolemia provided weak evidence for an association of this risk factor with Alzheimer's disease (`r obsHyperchol$AD$estimate`). A meta-analysis of MR analyses of genetic instruments that mimic statin use found weak evidence for an association of lipid lowering with Alzheimer's disease (`r mrStatin$AD$estimate`).

@@

Finally, only one of the three cohorts included in the IPD analysis (Whitehall II) contained data on Alzheimer's disease. However, a previously published analysis (Tynkkynen _et al._[@tynkkynen2018]) examining the association of lipid levels with Alzheimer's disease in this cohort was identified by the systematic review. The results of this previous analysis are incorporated into the results of the meta-analysis reported above, and so the IPD analysis provided no new information on this outcome.

@@

These causal questions were chosen for several reasons. Firstly, the work presented in the previous chapters had identified some evidence for an association in these exposure/outcome pairs. Secondly, given the long prodromal period between the onset of physiological changes in the brain and presentation of dementia symptoms, it seems likely that conditions during the mid-life period are particularly important. Finally, examining the causal impact of a lipid fraction on a specific outcome (e.g., Alzheimer's disease), rather than a composite outcome (e.g., all-cause dementia), focuses on a single causal pathway rather than including several conditions in the outcome definition that likely have very different mechanisms of disease. In addition, previous work presented in this thesis illustrated that the component conditions (in this case Alzheimer's disease and vascular dementia) are likely to be subject to very different confounding structures (see Section \@ref(cprd-confounding-by-ind)), and the use of a composite all-cause outcome in the triangulation framework may mask this.

@@

Once the causal question of interest had been defined, relevant individual results were obtained from the data sources described in Section \@ref(tri-data-sources). In a broad sense, this meant extracting results that examined the relationship between the exposure/outcome pair, either directly (e.g., non-randomised studies of LDL-c levels) or indirectly (e.g., non-randomised studies of statins, or MR studies using genetic instruments for LDL-c levels).

@@

&nbsp;<!----------------------------------------------------------------------->  
(ref:robLevelsMapping-caption) __Mapping of risk-of-bias judgements across assessment tools__ - The different acceptable levels of bias across the risk-of-bias assessment tools used were mapped to a harmonised set. Note that no mapping for "Critical" risk of bias was performed, as best practice recommends that studies at critical risk of bias are excluded from quantitative syntheses.

@@

@@

<!-- TODO Need to include some text here on examples of each: e.g., immortal time bias, or  -->

@@

One key feature of the domain-based risk-of-bias tools is that the domains are considered interchangeable - i.e., a "high" judgement in one domain is equivalent to a "high" judgement in any other. As such, for this analysis, I assumed that the distributions assigned to each extent of bias were identical across all values of $j$, that is, were identical for all domains in the risk-of-bias assessment tool. To illustrate for additive biases:

@@

(ref:priorsAdd-caption) __Prior distributions mapped to different extents of bias__ - Prior values assigned to additive and proportional biases, defined as log-normal distributions with mean $\mu$ and variance $\sigma^2$ using the notation $N(\mu,\sigma^{2})$. For additive biases, two scenarios were defined (see main text). Note that proportional biases are defined on the log scale for ease of calculation but are exponentiated prior to their use for adjustment (see Equations \@ref(eq:prop-total-mean) & \@ref(eq:prop-total-var) below).

@@

Where the direction of bias in a domain was unpredictable for a given result, the mean of the prior distribution for that domain was set to 0 but the appropriate variance for the recorded extent of bias was retained (e.g., for an unpredictable "High" additive bias, the distribution would be $N(0,0.1)$ under Scenario 1). In other words, adjusting for bias in a domain with an unpredictable direction of bias has no impact on the position of the effect estimate but does increase the uncertainty around it.

Under the  method reported in Turner _et al._,[@turner2009], total additive bias (denoted using subscript $\delta$) in result $i$ is then described by mean $\mu_{i\delta}$ and variance $\sigma_{i\delta}^{2}$, which are simply the sum of the mean and variance of all additive biases in the $i$th result. Similarly, the total proportional bias (denoted using subscript $\beta$) in the $i$th result is described by mean $\mu_{i\beta}$ and variance ${\sigma_{i\beta}}^2$. This are defined as

@@

where $\alpha_i$ and ${\psi_i}^2$  are the sum of mean and variance values for all proportional biases in the $i$th result.

@@

Using the method reported in Turner _et al._,[@turner2009] once total additive and proportional bias and indirectness are calculated, they are used to define an adjusted estimate for each result included in the synthesis.

@@

<!-- REVIEW YBS - I may have got confused but when you say favour the intervention I am assuming you mean a “protective effect” e.g., OR <1. But if there is differential misclassification so that statin use increases diagnosis of vascular dementia that would show a deleterious effect under the null hence shift to right e.g., OR >1. Perhaps I have misunderstood favour as < risk when in fact you mean > risk. Can you clarify?

@@

From the data sources identified in Section \@ref(tri-data-sources), I identified 7 results relevant to the causal effect of triglycerides at midlife on vascular dementia. Four were identified via the systematic review,`r tg_vad_citations` while Chapter \@ref(cprd-analysis-heading) provided two additional estimates from unanalysed datasets (CaPS and Whitehall II), and Chapter \@ref(ipd-heading) provides an estimate for the effect of fibrates, a treatment for hypertriglyceridemia. The extent, type and direction of predicted bias for each result can be seen in the bias-direction plot presented in Figure \@ref(fig:tgVadBiasDirection), stratified by study design.

@@

A further issue in this regard is that the rigour of the tools may not be equivalent. This issue can be illustrated by the Ostergaard _et al._ study[@ostergaard2015] having low risk of bias across all domains and thus receiving a substantial proportion of the weight in the first case study (Figure \@ref(fig:fpLdlAd)). While the RoB suite of tools (RoB 2.0, ROBINS-I, ROBINS-E) were developed by an expert working group and have been widely used, the Mamluk _et al._ tool[@mamluk2020] used to assess risk of bias in Mendelian randomisation studies was designed by the authors for use in their own review. If the Mamluk tool is failing to adequately assess bias in Mendelian randomisation studies, then any analysis based on its assessments is likely to be subject to bias. This observation is particularly problematic given the default position of the ROBINS-I and ROBINS-E tool to require a "Moderate" judgement in the confounding domain. Under these conditions, it is possible that a well-performed cohort study (low risk of bias across everything other that bias due to confounding) will be downweighted versus a potentially poor MR study.

@@

The core strength of this analysis is that it is based on a comprehensive systematic review, supplemented by two additional primary studies to fill in the identified evidence gaps. In general, triangulation of any sort should be considered a natural extension of evidence synthesis, and therefore should follow best practice in relation to finding and critically assessing all relevant information. Additionally, the approach presented here also builds on recent developments in bias assessment to "explode out" the component results of a meta-analysis and consider the effects of bias/indirectness in each separately.

@@

<!-- REVIEW Good to mention direct adjustment for bias, e.g., Greenlands writings -->

@@

* In this chapter, I narratively described the evidence identified or produced in previous chapters and highlighted the issues with a qualitative approach to triangulation when there are several sources of evidence.

@@

* I discussed the limitations of this proposed method, given the absence of empirically based priors for the impact of bias/indirectness, and proposed future research (i.e. meta-epidemiological studies) that would address them.

### 2021-12-16 08-Discussion.Rmd
End of day:

The proposed quantitative triangulation framework represents the final novel contribution of this thesis to evidence synthesis methodology. Using pre-specified adjustment distributions mapped to the results of the domain-based risk-of-bias and indirectness assessments, this approach provides a systematic way to account for biases/indirectness across studies. In addition, it explicitly seeks to include both direct and indirect evidence, and separates the assessment of bias/indirectness in each result from the definition of distributions representing the impact of bias/indirectness.

@@

Additionally, this thesis has also produced new evidence on the association of blood lipids with dementia outcomes in previously unanalysed cohorts, and provided additional information on a previously under-studied relationship (lipids/LRA and vascular dementia).

@@

As part of a reflective learning process throughout my PhD, I maintained a catalogue of failures (available in Appendix \@ref(appendix-catalogue-failures)) describing analytical mistakes, failed experiments, and unsuccessful grant applications. However, I found there was one central learning point from this thesis, namely to be slightly less ambitious when planning future research projects.

The proposal for the program of research underpinning this thesis was created in advance of starting my PhD because a detailed plan was required in order to secure the funding that supported me through my studies. In hindsight, the decision to take a broad approach to the inclusion of dementia outcomes resulted in a larger workload than anticipated, particularly when also considering evidence from multiple different study designs. Additionally, as dementia subtypes likely have very different aetiological pathways, it may have been better to focus on a single subtype (e.g. Alzheimer's disease) when considering the causal effects of lipids. Similarly, attempting to undertake a full individual participant data meta-analysis as a single part of a larger program of research was overly ambitious, given that data cleaning and harmonisation for just three cohorts turned out to be a substantial undertaking.

@@

This approach has already been used to assess long-term (11-20 years) safety outcomes of statin use, using participant recontact[@group2011] and linkage with electronic health records to identify events.[@ford2016] A similar approach could be used to assess the impact of randomisation to statins at midlife on subsequent risk of dementia outcomes. Use of data linkage would represent the most cost-efficient approach,[@llewellyn-bennett2018] though as noted in Chapter \@ref(cprd-analysis-heading), there is the potential for non-differential misclassification when defining dementia outcomes using electonic health records.

@@

As highlighted by the results of the systematic review presented in Chapter \@ref(sys-rev-results-heading), there is an absence of evidence on vascular dementia across the existing evidence base. While the primary studies presented here go some way towards supplementing this evidence base, future work is needed both to investigate the reasons for this evidence gap (e.g. due to an absence of primary data, challenges in the analysis of this outcome as documented in Chapter \@ref(cprd-analysis-heading), or a publication bias mechanism) and to address it.

@@

As noted in Chapter \@ref(sys-rev-results-heading), methods for the inclusion of Mendelian randomisation studies in evidence synthesis exercises are not yet sufficiently developed, particularly in comparison to other study designs. Future work could aim to validate search strategies for this study design, paying particular attention to the range of terms used to define the analytical approach (e.g. genetic instrumental variable analysis). In addition, as discussed in Section \@ref(tri-accuracy), the lack of an established risk-of-bias tool for Mendelian randomisation studies limits subsequent analyses that require systematic assessment of bias, such as the quantitative triangulation framework proposed here. Creation of a domain-based tool for this study design, which accounts for the differences in potential biases between one- and two-sample approaches, should be considered a priority.

>>>>

The proposed quantitative triangulation framework represents the final novel contribution of this thesis to evidence synthesis methodology. Using pre-specified adjustment distributions mapped to the results of the domain-based risk-of-bias and indirectness assessments, this approach provides a systematic way to account for biases/indirectness across studies. In addition, it explicitly seeks to include both direct and indirect evidence and separates the assessment of bias/indirectness in each result from the definition of distributions representing the impact of bias/indirectness.

@@

Additionally, this thesis has also produced new evidence on the association of blood lipids with dementia outcomes in previously unanalysed cohorts and provided additional information on a previously under-studied relationship (lipids/LRA and vascular dementia).

@@

As part of a reflective learning process throughout my PhD, I maintained a catalogue of failures (available in Appendix \@ref(appendix-catalogue-failures)) describing analytical mistakes, failed experiments, and unsuccessful grant applications. However, I found there was one central learning point from this thesis, namely, to be slightly less ambitious when planning future research projects.

The proposal for the program of research underpinning this thesis was created in advance of starting my PhD because a detailed plan was required in order to secure the funding that supported me through my studies. In hindsight, the decision to take a broad approach to the inclusion of dementia outcomes resulted in a larger workload than anticipated, particularly when also considering evidence from multiple different study designs. Additionally, as dementia subtypes likely have very different aetiological pathways, it may have been better to focus on a single subtype (e.g., Alzheimer's disease) when considering the causal effects of lipids. Similarly, attempting to undertake a full individual participant data meta-analysis as a single part of a larger program of research was overly ambitious, given that data cleaning and harmonisation for just three cohorts turned out to be a substantial undertaking.

@@

This approach has already been used to assess long-term (11-20 years) safety outcomes of statin use, using participant recontact[@group2011] and linkage with electronic health records to identify events.[@ford2016] A similar approach could be used to assess the impact of randomisation to statins at midlife on subsequent risk of dementia outcomes. Use of data linkage would represent the most cost-efficient approach,[@llewellyn-bennett2018] though as noted in Chapter \@ref(cprd-analysis-heading), there is the potential for non-differential misclassification when defining dementia outcomes using electronic health records.

@@

As highlighted by the results of the systematic review presented in Chapter \@ref(sys-rev-results-heading), there is an absence of evidence on vascular dementia across the existing evidence base. While the primary studies presented here go some way towards supplementing this evidence base, future work is needed both to investigate the reasons for this evidence gap (e.g., due to an absence of primary data, challenges in the analysis of this outcome as documented in Chapter \@ref(cprd-analysis-heading), or a publication bias mechanism) and to address it.

@@

As noted in Chapter \@ref(sys-rev-results-heading), methods for the inclusion of Mendelian randomisation studies in evidence synthesis exercises are not yet sufficiently developed, particularly in comparison to other study designs. Future work could aim to validate search strategies for this study design, paying particular attention to the range of terms used to define the analytical approach (e.g., genetic instrumental variable analysis). In addition, as discussed in Section \@ref(tri-accuracy), the lack of an established risk-of-bias tool for Mendelian randomisation studies limits subsequent analyses that require systematic assessment of bias, such as the quantitative triangulation framework proposed here. Creation of a domain-based tool for this study design, which accounts for the differences in potential biases between one- and two-sample approaches, should be considered a priority.

### 2021-12-16 front-and-back-matter/99-appendices.Rmd
End of day:

In addition to publications arising directly from this thesis (described in Section \@ref(literature-contributions) & included in Appendix \@ref(published-papers)), I have been involved in a number of external research projects. Taken together, these additional projects provided extensive experience in evidence synthesis methods and in collaborating across large teams.

During 2020, I was involved in the updating of the Preferred Reporting Items for Systematic reviews and Meta-Analyses (PRISMA) statement. My contributions were focused on the data sharing aspects of the guidelines and on building an interactive web application that would allow users to complete the PRISMA checklist. Publications arising from this work are listed below:

@@

medRxiv automatically links preprints to their published counterparts, and provides the DOI of the published article as part of the metadata on a given preprint. Using `medrxivr` to access this data, I could assess the two-year publication rate of preprints originally posted to medRxiv in July 2019. The code used for this analysis is presented below:

@@

* For Mendelian randomisation analyses, I had originally intended to use a panel of experts to assess he risk of bias in this study design. However, through private correspondence with the authors of a review of risk-of-bias assessment tools for Mendelian randomisation studies (since published[@spiga2021]), I became aware of the Mamluk _et al_ tool. Given this tool was considered to be the best available for assessing Mendelian randomisation studies and that it followed a domain-based approach, enabling comparison with tools for other study designs, I decided to use it in the review.

@@

* I had originally intended to assess sources of funding alongside bias. However, on further reading,[@sterne2013] I learned that any bias introduced by sources of funding would operate through one of the existing domains of the risk-of-bias tools. As such, I did not consider funding sources seperately.

 * Finally, I performed a dose response meta-analysis to assess the evidence for a non-linear realtionship between lipids and dementia outcomes. This marked a departure from the published protocol, as I had not expected the number of studies reporting dementia risk across lipid sub-categories to be sufficient for a dose response analysis.

@@

An overview of the number of hits per database, in addition to the full line-by-line search results for the each database, are presented below.

(ref:searchOverviewAppendix-caption) __Summary of number of records identified in each database searched__ - The full line-by-line search results for the each database, are presented in subsequent tables. Note there was substantial duplication across databases, resulting in several thousand duplicates being removed prior to title and abstract screening.

@@

(ref:minFU-cap) __Illustration of minimum follow-up period in two study designs__ - Panel A shows a study in which all participants enter the study and are assinged an exposure group at the same time, making it ease to define a minimum period of follow-up. In contrast, Panel B illustrates the time- varying analysis used in Chapter \@ref(cprd-analysis-heading), which makes it challenging to define an unbiased minimum period of follow-up.

@@

The paper describing the bias/indirectness-adjusted meta-analysis method used expert elicitation to define a distribution for  bias and indirectness across eight studies examining the effectiveness of routine anti-D prophylaxis in Rhesus negative women.[@turner2009] The authors consider five sources of bias (selection bias, performance bias, attrition bias, detection bias and other bias) and four sources of indirectness (population, intervention, control, outcome). For each domain of bias/indirectness, four experts were asked to define the 67% interval around the effect expected due to bias/indirectness in the absence of a true effect. Additive or multiplicative scales were used to capture the intervals for additive and proportional bias/indirectness, respectively. The authors then use assumed log-normality of these distributions to derive a mean and variance for each on the log scale.

The resulting data for two domains of bias (attrition and detection) in one study are shown in Table \@ref(tab:addTable1-table).

@@

I then took the average and maximum of the expert-averaged position and variance values. The results of this for the example data are presented in Table \@ref(tab:addTable3-table). 
&nbsp;<!----------------------------------------------------------------------->  
(ref:addTable3-caption) __Mean and maximum bias distributions, using example data__ - The mean and maximum of position and variance values desribed in Table \@ref(tab:addTable3-table) were calculated to give the average and maximum bias adjustment across studies and domains of bias.

@@

This two step approach (averaging across assessors before taking the average and maximum across domains/studies) was important. While the final average values will be the same whichever approach is taken, in a one-step process, the maximum value can be artificially inflated by a single expert's estimate. For example, in the illustrative data subset above, Assessor 2 assigns a distribution of N(-0.32,0.14), which would double the maximum position (0.16) and triple the maximum variance (0.0475) obtained when first averaging across experts.

@@

Reproducible and science has been a key theme running through this thesis, as reflected by the development of an open source tools to search preprinted literature, create visualisations of risk-of-bias data and to preform quantitative triangulation. In line with this, an open source copy of the code used to produce this thesis is available on GitHub, as is the code used to perform the analysis contained within it.

@@

<!-- \includepdf[pages=-,link=true,pagecommand={\thispagestyle{empty}}]{published/cprd_medrxiv.pdf} -->
<!-- \includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/medrxivr.pdf} -->
<!-- \includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/robvis.pdf} -->
<!-- \includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/DAS_comparison.pdf} -->
<!-- \includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/prevsci.pdf} -->

>>>>

In addition to publications arising directly from this thesis (described in Section \@ref(literature-contributions) and included in Appendix \@ref(published-papers)), I have been involved in a number of external research projects. Taken together, these additional projects provided extensive experience in evidence synthesis methods and in collaborating across large teams.

During 2020, I was involved in the updating of the Preferred Reporting Items for Systematic reviews and Meta-Analyses (PRISMA) statement. My contributions were focused on the data sharing aspects of the guidelines and on building the PRISMA2020 interactive web application. Publications arising from this work are listed below:

@@

medRxiv automatically links preprints to their published counterparts and provides the DOI of the published article as part of the metadata on a given preprint. Using `medrxivr` to access this data, I could assess the two-year publication rate of preprints originally posted to medRxiv in July 2019. The code used for this analysis is presented below:

@@

* For Mendelian randomisation analyses, I had originally intended to use a panel of experts to assess the risk of bias in this study design. However, through private correspondence with the authors of a review of risk-of-bias assessment tools for Mendelian randomisation studies (since published[@spiga2021]), I became aware of the Mamluk _et al_ tool. Given this tool was considered to be the best available for assessing Mendelian randomisation studies and that it followed a domain-based approach, enabling comparison with tools for other study designs, I decided to use it in the review.

@@

* I had originally intended to assess sources of funding alongside bias. However, on further reading,[@sterne2013] I learned that any bias introduced by sources of funding would operate through one of the existing domains of the risk-of-bias tools. As such, I did not consider funding sources separately.

 * Finally, I performed a dose response meta-analysis to assess the evidence for a non-linear relationship between lipids and dementia outcomes. This marked a departure from the published protocol, as I had not expected the number of studies reporting dementia risk across lipid sub-categories to be sufficient for a dose response analysis.

@@

An overview of the number of hits per database, in addition to the full line-by-line search results for each database, are presented below.

(ref:searchOverviewAppendix-caption) __Summary of number of records identified in each database searched__ - The full line-by-line search results for each database, are presented in subsequent tables. Note there was substantial duplication across databases, resulting in several thousand duplicates being removed prior to title and abstract screening.

@@

(ref:minFU-cap) __Illustration of minimum follow-up period in two study designs__ - Panel A shows a study in which all participants enter the study and are assigned an exposure group at the same time, making it easy to define a minimum period of follow-up. In contrast, Panel B illustrates the time- varying analysis used in Chapter \@ref(cprd-analysis-heading), which makes it challenging to define an unbiased minimum period of follow-up.

@@

The paper describing the bias/indirectness-adjusted meta-analysis method used expert elicitation to define a distribution for  bias and indirectness across eight studies examining the effectiveness of routine anti-D prophylaxis in Rhesus negative women.[@turner2009] The authors consider five sources of bias (selection bias, performance bias, attrition bias, detection bias and other bias) and four sources of indirectness (population, intervention, control, outcome). For each domain of bias/indirectness, four experts were asked to define the 67% interval around the effect expected due to bias/indirectness in the absence of a true effect. Additive or multiplicative scales were used to capture the intervals for additive and proportional bias/indirectness, respectively. The authors then derived a mean and variance for each on the log scale. The resulting data for two domains of bias (attrition and detection) in one study are shown in Table \@ref(tab:addTable1-table).

@@

I then took the average and maximum of the expert-averaged position and variance values. The results of this for the example data are presented in Table \@ref(tab:addTable3-table). 
&nbsp;<!----------------------------------------------------------------------->  
(ref:addTable3-caption) __Mean and maximum bias distributions, using example data__ - The mean and maximum of position and variance values described in Table \@ref(tab:addTable3-table) were calculated to give the average and maximum bias adjustment across studies and domains of bias.

@@

This two-step approach (averaging across assessors before taking the average and maximum across domains/studies) was important. While the final average values will be the same whichever approach is taken, in a one-step process, the maximum value can be artificially inflated by a single expert's estimate. For example, in the illustrative data subset above, Assessor 2 assigns a distribution of N(-0.32,0.14), which would double the maximum position (0.16) and triple the maximum variance (0.0475) obtained when first averaging across experts.

@@

Reproducible and science has been a key theme running through this thesis, as reflected by the development of an open-source tools to search preprinted literature, create visualisations of risk-of-bias data and to preform quantitative triangulation. In line with this, an open-source copy of the code used to produce this thesis is available on GitHub, as is the code used to perform the analysis contained within it.

@@

\includepdf[pages=-,link=true,pagecommand={\thispagestyle{empty}}]{published/cprd_medrxiv.pdf}
\includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/medrxivr.pdf}
\includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/robvis.pdf}
\includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/DAS_comparison.pdf}
\includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/prevsci.pdf}

### 2021-12-16 front-and-back-matter/_abstract.Rmd
End of day:

<!-- TODO This needs to be less than 300 words! -->

__Background:__ Identification of causal relationships between modifiable targets and risk of dementia is central to the development of evidence-based prevention strategies. As such, this thesis sought to triangulate the causal effects of lipid levels (total cholesterol, LDL-c, HDL-c, and triglycerides) on risk of incident dementia (all-cause dementia, Alzheimer's disease, and vascular dementia) using existing and new evidence.

@@

__Results:__ The systematic review did not identify a consistent effect of blood lipids on any dementia outcome across study designs considered, though there was some suggestion of a protective effect of LDL-c lowering on all-cause dementia and Alzheimer's disease in observational studies of statin use. Analysis of the CPRD data provided weak evidence for a protective effect of statins on all-cause dementia and Alzheimer's disease but suggested a harmful association with vascular dementia,  likely due to confounding by indication. The IPD meta-analysis provided some evidence for only the association of triglycerides and vascular dementia. Finally, the quantitative triangulation analysis did not provide strong evidence for the causal effect of blood lipids on dementia outcomes.

>>>>

__Background:__ Identification of causal relationships between modifiable risk factors and incidence of dementia is central to the development of evidence-based prevention strategies. As such, this thesis sought to triangulate the causal effects of lipid levels (total cholesterol, LDL-c, HDL-c, and triglycerides) on risk of incident dementia (all-cause dementia, Alzheimer's disease, and vascular dementia) using existing and new evidence.

@@

__Results:__ The systematic review did not identify a consistent effect of blood lipids on any dementia outcome across study designs considered, though there was some suggestion of a protective effect of LDL-c lowering on all-cause dementia and Alzheimer's disease in observational studies of statin use. Analysis of the CPRD data provided weak evidence for a protective effect of statins on all-cause dementia and Alzheimer's disease but suggested a harmful association with vascular dementia,  likely due to confounding by indication. The IPD meta-analysis provided some evidence for the association of triglycerides and vascular dementia only. Finally, the quantitative triangulation analysis did not provide strong evidence for the causal effect of blood lipids on dementia outcomes.
